---
title: "Final_Lab_SG"
author: "Sarah Grant"
format: 
  html:
    message: false
    echo: false
---

```{r}
#| message: false
#| warning: false
here::i_am("science_fiction.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(stringr)
library(tidyverse)
```

```{r}
#| message: false
#| warning: false
planets <- vroom(here("data/data_planets.csv"))
stardata <- vroom(here("data/star_data.csv"))
startype <- vroom(here("data/star_types.csv"))
yerkes <- vroom(here("data/Yerkes_luminosity_classes.csv"))
```

## Question 1

Commit of main dataset complete

## Question 2 - Link to Github Repository

Please find the github project via this [link](https://github.com/sarahgrant1/FinalLab).
Please note that this is named differently to the recommended name science_lab. This was due to me creating this repository before recieving the email with the correct name so the commits are sent to this Github repository. Sorry about any confusion!

## Question 3 - Verifying the Validity of the Loading Process

```{r}
expected_planets = 1085
expected_variables = 8
```

```{r}
if (nrow(planets) == expected_planets && ncol(planets) == expected_variables) {
  cat("Loading process has been successful. Number of planets and variables indeed match the expected values.\n")
} else {
  stop("Uh oh! Loading process has not been successful.\n")
}
```

```{r}
planet <- na.omit(planets)
numerical_variables <- c("Mass (Earth)", "Relative Radius", "density (earth)")
non_numerical_variables <- sapply(planets[, numerical_variables], function(x) any(!is.na(as.numeric(x)) & !is.finite(as.numeric(x))))

if (any(non_numerical_variables)) {
  stop("Uh oh! Some of the values within these variables are not recognised as numerical.\n")
} else {
  cat("All numerical values are recognised as numerical.")
}
```

## Question 4

```{r}
num_observations <- nrow(planets)
num_variables <- ncol(planets)

missing_values_per_var <- colSums(is.na(planets))

missing_obvs <- sum(apply(planets, 1, anyNA))

table_qu4 <- data.frame(
  Metric = c("Number of Observations", "Number of Variables", "Missing Values per Variable", "Observations with at Least One Missing Value"),
  Value = c(num_observations, num_variables, toString(missing_values_per_var), missing_obvs)
)

knitr::kable(table_qu4, format = "markdown")
```

## Question 5

```{r}
#| message: false
#| warning: false
ggplot(planets, aes(x= `Mass (Earth)`, y = `density (earth)`)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Density of a Planet as a Funtion of its Mass",
       x = "Mass (Earth)",
       y = "Density (Earth)") +
  theme_minimal()
```

## Question 6

```{r}
densest_planets <- planets %>%
  arrange(desc(`density (earth)`)) %>%
  slice_head(n=5)

least_dense_planets <- planets %>%
  arrange(`density (earth)`) %>%
  slice_head(n=5)

combined_extreme_planets <- bind_rows(
  mutate(select(densest_planets, P_name, `density (earth)`), Extreme_Type = "Densest"),
  mutate(select(least_dense_planets, P_name, `density (earth)`), Extreme_Type = "Least Dense")
)

kable(combined_extreme_planets, format = "markdown", caption = "Top 5 Densest and Top 5 Least Dense Planets")
```

## Question 7
```{r}
split_and_reframe <- function(planets, variable_name, new_variable_name) {
  planets %>%
    select(planet_idx, {{variable_name}}) %>%
    mutate(str_split({{variable_name}}, pattern = ";", simplify = TRUE), .list = new_variable_name) %>%
    pull(planets)
}

```
## Question 8
